Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
some edits

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation
$s_1 \approx s_2$, where $s_1 \in \effCEK{C}$ and $s_2 \in \ineffCEKD{}$.
It is the smallest congruence relation that also includes the
pairs induced by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation. This handles the way in which
\effCEK{C} uses $wrap$ to push identity coercions onto continuations
whereas \ineffCEKD{} does not.
%
We remark that this bisimulation relation is unusual in that it uses
some ADT operations in its definition, such as $id$, $cast$, and
$seq$, which is necessary because \effCEK{C} and the bisimulation
is parameterized over the Cast ADT.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending the right-hand
side with a cast $T_1 \Rightarrow^l T_2$ and sequencing
$cast(T_1,l,T_2)$ with $c$ on the left-hand side, produces related
continuations. This takes care of the difference between
the transition rules for $e \langle T_1 \Rightarrow^l T_2$
on the two machines.
%
If two values of type $I$ are related, injecting them to $\TOOdyn$
gives related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a closure in \ineffCEKD, the closure in \effCEK{C} must
have identities as its casts.
%
If a closure on the left is related to a value, we can apply
equivalent casts on both sides to obtain related values.
%
The rules for products and sums are similar.

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}}
  \end{gather*}
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
  \\ \\
   %% This one is an obvious congruence rule. -Jeremy
   %% \inference{}{
   %%  \hcvOOtt \approx \vOOtt : \POOunit
   %% }
   %% \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

%\todo[inline]{Give the motivation for the following lemma. Where
%  does it get used? As in, which case of which other lemma or theorem? -JS
%\\---\\
%Partially fixed. Will say in which cases after finishing the bisim proof. -KC}

We use the following lemma in our bisimulation proof when cast applications 
happen.

@@

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $\judgeSreduce{C}{s_1}{s_3}$ then
	$\judgeDreduceTrans{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
	\item If $\judgeDreduce{s_2}{s_4}$ then
	$\judgeSreduceTrans{C}{s_1}{s_3}$,
	$\judgeDreduceTrans{s_4}{s_5}$, and
	$s_3 \approx s_5$ for some $s_3$ and $s_5$.
	\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
%\begin{enumerate}
%  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
%  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
%    then $\judgeDreduce{s_2}{s_4}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_4$
%  \item If $\judgeDreduce{s_2}{s_4}$
%    then $\judgeSreduce{C}{s_1}{s_3}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_3$
%\end{enumerate}
%{\color{red}
%  Jeremy's attempt to reformulate in a more standard way that
%  remains equivalent:
%\begin{enumerate}
%\item If $\judgeSreduce{C}{s_1}{s_3}$ then
%  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
%\item If $\judgeDreduce{s_2}{s_4}$ then
%  $\judgeSreduce{C}{s_1}{s_3}$,
%  $\judgeDreduceTrans{s_4}{s_5}$, and
%   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
%\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%\end{enumerate}
%}
%Or diagrammatically, 
%  \[\begin{array}{clclc}
%  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%  \rotatebox[origin=c]{90}{$\approx$} 
%  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%  s_{2} & 
%  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
%  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follows from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

  By looking at the definition of $v \approx v$, we know
  \[ c_{11} = seq(seq(id(T_1),cast(T_1,l,T_3)) \dots) \]
  \[ c_{12} = seq(seq(id(T_2),cast(T_2,l,T_4)) \dots) \] 
  \[ v_2 = \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{
    \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
  \]
  Thus we have the following reduction in \effCEK{C}. The equation 
  reasoning employs the premise that $C$ is a monoid.
  \[
  \begin{array}{ll}
&
\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{wrap(\kOOcar{\kappa})}
\\
\longmapsto_\text{\effCEK{C}} &
\sOOreturn{v_{11}}{ext(c_{11},\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(seq(seq(id(T_1),cast(T_1,l,T_3))\dots),\kappa)}
\\
= &
\sOOreturn{v_{11}}{
	ext(id(T_1),ext(cast(T_1,l,T_3),ext(\dots,\kappa)))
}
\\
= &
\sOOreturn{v_{11}}{ext(cast(T_1,l,T_3),ext(\dots,\kappa)}
  \end{array}
  \]

And we have the following reduction in \ineffCEKD, where the last state is 
related to the above \effCEK{C} state.

\[
\begin{array}{ll}
& \sOOreturn{
  \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{   
  \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
}{\kOOcar{\kappa}}
\\
\longmapsto_{\mathcal{D}}^{*} &
\sOOreturn{(\vOOcons{v_{21}}{v_{22}})}{
  \kOOcar{
  \kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}}
\\
\longmapsto_{\mathcal{D}} &
\sOOreturn{v_{21}}{
  \kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}
\end{array}
\]

Other cases are similar. Of course, we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case splitting, we employ $cast(T,l,T)=id(T)$ to relate the continuation 
functions.
\end{proof}

>>>>

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation
$s_1 \approx s_2$, where $s_1 \in \effCEK{C}$ and $s_2 \in \ineffCEKD{}$.
It is the smallest congruence relation that also includes the
pairs induced by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation. This handles the way in which
\effCEK{C} uses $wrap$ to push identity coercions onto continuations
whereas \ineffCEKD{} does not.
%
We remark that this bisimulation relation is unusual in that it uses
some ADT operations in its definition, such as $id$, $cast$, and
$seq$, which is necessary because \effCEK{C} and the bisimulation
is parameterized over the Cast ADT.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending the right-hand
side with a cast $T_1 \Rightarrow^l T_2$ and sequencing
$cast(T_1,l,T_2)$ with $c$ on the left-hand side, produces related
continuations. This takes care of the difference between
the transition rules for $e \langle T_1 \Rightarrow^l T_2$
on the two machines.

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}} \\
  \end{gather*} 
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
  \\[1ex]
   %% This one is an obvious congruence rule. -Jeremy
   %% \inference{}{
   %%  \hcvOOtt \approx \vOOtt : \POOunit
   %% }
   %% \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\[1ex]
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\[1ex]
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2}}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\begin{array}{l}
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
\end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\[1ex]
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2}
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
  \begin{array}{l}
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  \end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\[1ex]
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

The bisimulation relation for values is designed to make
Lemma~\ref{thm:lem-apply-cast-SD} true, that is, if $v_1 \approx v_2$,
then
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
So if two values of type $I$ are related, injecting them to $\TOOdyn$
gives related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a closure in \ineffCEKD, the closure in \effCEK{C} must
have identities as its casts.
%
If a closure on the left is related to a value, we can apply
equivalent casts on both sides to obtain related values.
%
The rules for products and sums are similar.
%
Lemma~\ref{thm:lem-apply-cast-SD} is used in our bisimulation proof to
handle the transitions that apply casts to values.

@@

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $\judgeSreduce{C}{s_1}{s_3}$ then
	$\judgeDreduceTrans{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
	\item If $\judgeDreduce{s_2}{s_4}$ then
	$\judgeSreduceTrans{C}{s_1}{s_3}$,
	$\judgeDreduceTrans{s_4}{s_5}$, and
	$s_3 \approx s_5$ for some $s_3$ and $s_5$.
	\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
\end{enumerate}
%\begin{enumerate}
%  \item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%%  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
%  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
%    then $\judgeDreduce{s_2}{s_4}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_4$
%  \item If $\judgeDreduce{s_2}{s_4}$
%    then $\judgeSreduce{C}{s_1}{s_3}$ 
%    and  $s_3 \approx^{*} s_4$
%    for some $s_3$
%\end{enumerate}
%{\color{red}
%  Jeremy's attempt to reformulate in a more standard way that
%  remains equivalent:
%\begin{enumerate}
%\item If $\judgeSreduce{C}{s_1}{s_3}$ then
%  $\judgeDreducePlus{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$.
%\item If $\judgeDreduce{s_2}{s_4}$ then
%  $\judgeSreduce{C}{s_1}{s_3}$,
%  $\judgeDreduceTrans{s_4}{s_5}$, and
%   $s_3 \approx s_5$ for some $s_3$ and $s_5$.
%\item $s_1 = \sOOhalt{o}$ if and only if  $s_2 = \sOOhalt{o}$
%\end{enumerate}
%}
%Or diagrammatically, 
%  \[\begin{array}{clclc}
%  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%  \rotatebox[origin=c]{90}{$\approx$} 
%  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%  s_{2} & 
%  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
%  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) follows from the definition of $s \approx s$.
  Part (2) and (3) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to prove that 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

  By looking at the definition of $v \approx v$, we know
  \begin{align*}
   c_{11} &= seq(seq(id(T_1),cast(T_1,l,T_3)) \dots) \\
   c_{12} &= seq(seq(id(T_2),cast(T_2,l,T_4)) \dots) \\
   v_2 &= \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{
    \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
  \end{align*}
  Thus we have the following reduction in \effCEK{C}. Then we employ
  equational reasoning that relies on the premise that $C$ is a
  monoid.
  \[
  \begin{array}{ll}
&
\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{wrap(\kOOcar{\kappa})}
\\
\longmapsto_\text{\effCEK{C}} &
\sOOreturn{v_{11}}{ext(c_{11},\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(seq(seq(id(T_1),cast(T_1,l,T_3))\dots),\kappa)}
\\
= &
\sOOreturn{v_{11}}{
	ext(id(T_1),ext(cast(T_1,l,T_3),ext(\dots,\kappa)))
}
\\
= &
\sOOreturn{v_{11}}{ext(cast(T_1,l,T_3),ext(\dots,\kappa)}
  \end{array}
  \]

We have the following corresponding reduction in \ineffCEKD, where the
last state is related to the above \effCEK{C} state.
\[
\begin{array}{ll}
& \sOOreturn{
  \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{   
  \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
}{\kOOcar{\kappa}}
\\
\longmapsto_{\mathcal{D}}^{*} &
\sOOreturn{(\vOOcons{v_{21}}{v_{22}})}{
  \kOOcar{
  \kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}}
\\
\longmapsto_{\mathcal{D}} &
\sOOreturn{v_{21}}{
  \kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}
\end{array}
\]

Other cases are similar. Of course, we need the lemma about
$applyCast$ (Lemma~\ref{thm:lem-apply-cast-SD}) when cast applications
happen in both machines.  And in case splitting, we employ
$cast(T,l,T)=id(T)$ to relate the continuation functions.
\end{proof}
